PyPy: raising the level for Dynamic Language implementations

### 2007-05-27 talk/dyla2007/dyla.tex
okok. revert the last one since armin's latex doesn't seem to like it.
a quick try describing PyPy's approach without contrasting it to .NET/CLI too much - rather focusing on presenting PyPy as the next evolutional step to raise the level of language implementations, and still taking advantage of existing new approaches/VMs.

\documentclass[utf8x]{llncs}

\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}

@@

\title{How to Not write Virtual Machines for Dynamic Languages}

@@

Dynamic languages are traditionally implemented by writing a virtual
machine for them, centered around an interpreter and/or a built-in
compiler and providing the object model and memory management. When
a language becomes more successful, the limitations of such an implementation
lead to the emergence of alternative implementations that try to solve some
of the problems. Another reason for new implementations is the desire to have
the language integrate well with existing, well-tuned object-oriented virtual
machine like Java's. In this paper, we describe the mechanisms that lead to an
abundance of implementations, and explore some of the limitations of standard
VMs.  We propose a different complementary alternative: flexibly generating
virtual machines instead
of writing them by hand.  This approach has been validated by the PyPy
project, with which we can automatically insert many features into the
generated virtual machines -- including good just-in-time compilers tuned
to the dynamic language at hand.
\footnote{This research is partially supported by the EU funded %research
 project:
IST 004779 PyPy (PyPy: Implementing Python in Python).} \\

>>>>

\documentclass{llncs}

\usepackage{makeidx}
\usepackage{graphicx}

@@

\title{PyPy: raising the level for Dynamic Language implementations}

@@

Dynamic languages are usually implemented in lower level environments, 
traditionally in C/Posix.  Recently .NET and the JVM are emerging as 
new higher-level environments, offering more built-in support for 
memory management, threading models, dynamic optimization techniques
and also offering an abundance of libraries for application development.   
However, manually implementing languages for a standardized runtime environment
may not be the last word on bringing dynamic language implementations to a higher level. 
The PyPy project successfully experimented with an approach that 
uses extensive meta-programming techniques to analyse and transform the 
single source of an language "specification" and automatically adds
lower level and runtime-dependent aspects, successfully producing 
efficent virtual machines for an runtime environment of choice. 
The meta-programming facilities are largely re-useable for implementing 
different computer languages. We thus argue in this paper that 
one should not write interpreters manually and dependent on lower 
level and runtime details - but rather explore and exploit 
meta-programming facilities which themselves can be written 
and tested in a very high language, overall raising the level 
of implementing dynamic languages.  While there is considerable 
complexity liying in such a meta-programming tool chain, it is 
re-usable for different language implementations and generally
helps to avoid having to implement language semantics and 
optimizations redundantly.

\footnote{This research is partially supported by the EU funded %research
 project:
IST 004779 PyPy (PyPy: Implementing Python in Python).} \\
